\documentclass[11pt,a4paper,sans]{moderncv} 


\moderncvstyle{banking}                             % style options are 'casual' (default), 'classic', 'banking', 'oldstyle' and 'fancy'
\moderncvcolor{blue}                               % color options 'black', 'blue' (default), 'burgundy', 'green', 'grey', 'orange', 'purple' and 'red'

% Last updated footnote
\usepackage{fancyhdr}
\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\cfoot{Last updated \today.}

% adjust the page margins
\usepackage[scale=0.75]{geometry}
\usepackage{enumitem}

% personal data
\name{R. Teal}{Witter}
\email{rtealwitter@nyu.edu}
\homepage{www.rtealwitter.com}
\social[github]{rtealwitter}


\begin{document}

\makecvtitle

\section{Education}
\cventry{September 2020--Present}{PhD in Computer Science}{New York University}{New York, NY}{}{Advised by \href{https://cse.engineering.nyu.edu/~hstein/}{Lisa Hellerstein} and \href{https://www.chrismusco.com/}{Christopher Musco}}  % arguments 3 to 6 can be left empty
\cventry{February 2017--May 2020}{BA in Mathematics, BA in Computer Science}{Middlebury College}{Middlebury, VT}{}{Phi Beta Kappa, Summa Cum Laude}

\section{Research Interests}
Algorithms for Social Good $\bullet$ Explainable AI  $\bullet$ Fairness $\bullet$ Randomized Linear Algebra $\bullet$ Machine Learning $\bullet$ Deep Learning $\bullet$ Discrete Optimization $\bullet$ Graph Theory $\bullet$ Quantum Computing

\section{National Awards}

\cvitemwithcomment{}{NSF Graduate Research Fellow}{2022-2025}
\cvitemwithcomment{}{Goldwater Scholar}{2019}
\cvitemwithcomment{}{Academic All-American}{2015}

\section{Teaching}
\cventry{Winter 2024}{Course Instructor}{Randomized Algorithms for Data Science}{Middlebury CSCI 1052}{}{}
\cventry{Winter 2023}{Course Instructor}{Deep Learning}{Middlebury CSCI 1051}{}{}
\cventry{Fall 2022, Spring 2023, Fall 2023}{Course Assistant}{Deep Learning}{NYU CS-GY 6953}{}{}
\cventry{Fall 2021, Spring 2022, Fall 2023}{Course Assistant}{Algorithmic Machine Learning and Data Science}{NYU CS-GY 6763}{}{}
\cventry{Spring 2021, Spring 2023}{Course Assistant}{Machine Learning}{NYU CS-GY 6923}{}{}

% Publications from a BibTeX file using the biblatex package
\section{Preprints}
\textit{In the tradition of theoretical computer science, authors appear in alphabetical order unless otherwise marked with an asterisk.}

\begin{enumerate}[label={[\arabic*]},align=left]
    \item Y. Liu, R. T. Witter, F. Korn, T. Alrashed, D. Paparas, J. Freire. \textit{Kernel Banzhaf: A Fast and Robust Estimator for Banzhaf Values}. 2024.*
    \item C. Musco, R. T. Witter. \textit{Leverage SHAP: Estimating Shapley Values with Leverage Score Sampling}. 2024.*
    \item K. Arabi, B. Feuer, R. T. Witter, C. Hegde, N. Cohen. \textit{Hidden in the Noise: Two-Stage Robust Watermarking for Images}. 2024.*
    \item R. T. Witter, L. Rosenblatt. \textit{FairlyUncertain: A Comprehensive Evaluation of Uncertainty in Algorithmic Fairness}. 2024.*
    \item R. T. Witter, L. Hellerstein. \textit{Minimizing Cost Rather Than Maximizing Reward in Restless Multi-Armed Bandits}. 2024.*
\end{enumerate}


\section{Peer-Reviewed Publications}

\begin{enumerate}[label={[\arabic*]},align=left]
    \setcounter{enumi}{5}
    \item R. T. Witter and C. Musco. \textit{Benchmarking Estimators for Natural Experiments: A Novel Dataset and a Doubly Robust Algorithm}. Conference on Neural Information Processing Systems, 2024.*
    \item R. T. Witter and  L. Rosenblatt. \textit{I Open at the Close: A Deep Reinforcement Learning Evaluation of Open Streets Initiatives}. AAAI Conference on Artificial Intelligence, 2024.*
    \item M. Czekanski, S. Kimmel, R. T. Witter. \textit{Robust and Space-Efficient Dual Adversary Quantum Query Algorithms}. European Symposium on Algorithms, 2023.
    \item L. Rosenblatt, R. T. Witter. \textit{Counterfactual Fairness Is Basically Demographic Parity}. AAAI Conference on Artificial Intelligence, 2023.*
    \item L. Hellerstein, D. Kletenik, N. Liu, R. T. Witter. \textit{Adaptivity Gaps for the Stochastic Boolean Function Evaluation Problem}. Workshop on Approximation and Online Algorithms, 2022.
    \item L. Hellerstein, T. Lidbetter, R. T. Witter. \textit{A Local Search Algorithm for the Min-Sum Submodular Cover Problem}. International Symposium on Algorithms and Computation, 2022.
    \item C. Musco, I. Ramesh, J. Ugander, R. T. Witter. \textit{How to Quantify Polarization in Models of Opinion Dynamics}. International Workshop on Mining and Learning with Graphs, 2022.
    \item S. Kimmel, R. T. Witter. \textit{A Query-Efficient Quantum Algorithm for Maximum Matching on General Graphs}. Algorithms and Data Structures Symposium, 2021.
    \item R. T. Witter. \textit{Backgammon is Hard}. International Conference on Combinatorial Optimization and Applications, 2021.
    \item R. T. Witter, A. Lyford. \textit{Applications of Graph Theory and Probability in the Board Game Ticket to Ride}. International Conference on the Foundations of Digital Games, 2020.*
    \item K. DeLorenzo, S. Kimmel, R. T. Witter. \textit{Applications of the Quantum Algorithm for st-Connectivity}. Conference on the Theory of Quantum Computation, Communication and Cryptography, 2019.
\end{enumerate}

\section{Talks}
\textbf{}

\textbf{Estimating the Impact of Social Programs in Resource-Constrained Settings}

\cvitemwithcomment{}{NYU-KAIST Inclusive AI Workshop}{November 2023}

\textbf{Robust and Space-Efficient Dual Adversary Quantum Query Algorithms}

\cvitemwithcomment{}{Centrum Wiskunde \& Informatica QuSoft Seminar}{September 2023}

\cvitemwithcomment{}{Quantum Computing and Optimization Minisymposium at SIAM NNP}{October 2023}

\textbf{Adaptivity Gaps for the Stochastic Boolean Function Evaluation Problem}

\cvitemwithcomment{}{Workshop on Approximation and Online Algorithms}{September 2022}

\textbf{How to Quantify Polarization in Models of Opinion Dynamics}

\cvitemwithcomment{}{International Workshop on Mining and Learning with Graphs}{August 2022}

\textbf{A Local Search Algorithm for the Min-Sum Submodular Cover Problem}

\cvitemwithcomment{}{International Symposium on Algorithms and Computation}{December 2022}

\cvitemwithcomment{}{International Workshop on Mining and Learning with Graphs}{January 2022}

\textbf{Backgammon is Hard}

\cvitemwithcomment{}{International Workshop on Mining and Learning with Graphs}{December 2021}

\textbf{A Query-Efficient Quantum Algorithm for Maximum Matching on General Graphs}

\cvitemwithcomment{}{International Workshop on Mining and Learning with Graphs}{August 2021}

\textbf{Applications of Graph Theory and Probability in the Board Game \textit{Ticket to Ride}}

\cvitemwithcomment{}{International Workshop on Mining and Learning with Graphs}{September 2020}

\cvitemwithcomment{}{Contributed Paper Session at the Joint Mathematics Meetings}{January 2020}

\textbf{Applications of the Quantum Algorithm for \textit{st}-Connectivity}

\cvitemwithcomment{}{Conference on the Theory of Quantum Computation, Communication and Cryptography}{June 2019}

\section{Service}

\textbf{Conference Reviewing}

QIP 2022, ICALP 2022, TQC 2022, NeurIPS 2023, ICLR 2024, ICML 2024, NeurIPS 2024, AAAI~2025

\textbf{Journal Reviewing}

Information Processing Letters, Theoretical Computer Science

\section{Outreach}

\cventry{Spring 2021-2023}{Instructor}{Extracurricular Coding Club}{Brooklyn International High School}{}{}

\section{Advising}

\cventry{Summer 2024}{Barnard College '26}{Syna Sachdeva}{Gaussian Splatting with Latent Representations}{}{}

\cventry{Spring 2024-Present}{New York University '25}{Jack Liu}{Latent Guidance of Large Language Models}{}{}

\cventry{Summer 2022}{Brooklyn International High School '22}{Xiaorui Lei}{Active Learning and Importance Sampling}{}{}

\cventry{Summer 2022}{Brooklyn International High School '22}{Bryant Chen}{Active Learning and Importance Sampling}{}{}

\section{References}

\textbf{Christopher Musco}

Assistant Professor of Computer Science and Engineering, New York University

cmusco@nyu.edu

\vspace{1em}

\textbf{Lisa Hellerstein}

Professor of Computer Science and Engineering, New York University

lisa.hellerstein@nyu.edu

\vspace{1em}

\textbf{Shelby Kimmel}

Associate Professor of Computer Science, Middlebury College

skimmel@middlebury.edu

\end{document}
